### 2020-05-27 manuscript.tex
Start adding EPC-RK4 and SIE integrators

\documentclass[3p,authoryear]{elsarticle}
\journal{Annals of Nuclear Energy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[charter]{mathdesign} % Type-1 Latin Modern modern
\usepackage[T1]{fontenc}         % Use T1 encoding instead of OT1
\usepackage[utf8]{inputenc}      % Use UTF8 input encoding
\usepackage{microtype}           % Improve typography
\usepackage{booktabs}            % Publication quality tables
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage{multirow}

\sisetup{list-final-separator = {, and }}

% Captions for figures and tables
\usepackage[labelfont=bf]{caption}
\captionsetup[figure]{labelsep=period, name=Fig.}
\captionsetup[table]{labelsep=newline}

\usepackage{color}
\definecolor{aneblue}{RGB}{0,128,173}
\usepackage[colorlinks,breaklinks,bookmarksnumbered,bookmarksopen]{hyperref}
\AtBeginDocument{
  \hypersetup{linkcolor=aneblue, citecolor=aneblue, urlcolor=aneblue}
}

\usepackage[capitalise,nameinlink]{cleveref}

\newtheorem{lemma}{Lemma}

\newcommand{\vect}[1]{\mathbf{#1}} % vectors and matrices

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Depletion capabilities in the OpenMC Monte Carlo particle transport code}

\author[anl]{Paul K. Romano\corref{cor1}}
\ead{promano@anl.gov}
\cortext[cor1]{Corresponding author. Tel.: +1 630 252 6779.}

\author[lanl]{Colin Josey}
\ead{cjosey@lanl.gov}

\author[gatech]{Andrew E. Johnson}
\ead{dasindrew@gatech.edu}

\author[tsinghua]{Jingang Liang}
\ead{jingang@tsinghua.edu.cn}

\address[anl]{Argonne National Laboratory, 9700 S. Cass Ave, Lemont, IL 60439, United States}
\address[lanl]{Los Alamos National Laboratory, P.O. Box 1663, Los Alamos, NM 87545, United States}
\address[gatech]{Georgia Institute of Technology, 770 State St NW, Atlanta, GA 30318, United States}
\address[tsinghua]{Institute of Nuclear and New Energy Technology, Tsinghua University, Beijing, China}

\begin{abstract}

\end{abstract}

\begin{keyword}
  Monte Carlo, depletion
\end{keyword}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

When a material in a system is subject to irradiation over a long period of
time, nuclides within the material will transmute due to nuclear reactions and
spontaneous radioactive decay. The time-dependent process by which nuclides
transmute under irradiation is known as \emph{depletion}, \emph{burnup}, or
\emph{transmutation}. Modeling this process plays an important role in the
design and licensing of nuclear reactors~\citep{betzler2019ned}. Consequently,
many codes have been developed that solve the underlying equations for the
change in material compositions over time.

When modeling nuclear reactors, depletion is inherently linked with solving the
neutron transport equation. Nuclear reaction rates, which determine the rate at
which nuclides are transmuted, must be provided from a transport code.
Conversely, the change in material composition directly affects the solution of
the transport equation. As will be shown in \cref{sec:methods}, this results in
a non-linear coupling between transport and depletion. The implication is that
depletion codes are rarely used in isolation---normally, they are directly
coupled with a neutron transport code.

With continued advances in computing, Monte Carlo (MC) particle transport
simulations are now widely used in reactor design and analysis in a variety of
roles. The earliest efforts to couple MC particle transport with depletion date
back to the 1990s~\citep{moore1995inel,trellue1998lanl}. These early efforts
involved taking existing MC and depletion codes, such as
MCNP~\citep{goorley2012nt} and ORIGEN~\citep{croff1983nt}, and putting a
``wrapper'' around them that orchestrated the transfer of solutions from one
code to the other. Thus, the actual transport/depletion codes themselves
required no source code changes. In the last decade, many MC code developers
have chosen to directly include a depletion solver as a first-rate feature,
including MCNPX~\citep{waters2007aip} (which has since been merged into
MCNP6~\citep{goorley2012nt}), Serpent~\citep{leppanen2015ane}, and
MC21~\citep{griesheimer2015ane}.

In the present work, we describe a new depletion solver in
OpenMC~\citep{romano2015ane1}, an open source, community-developed MC particle
transport code. To date, OpenMC has not had a built-in depletion solver. This
has led to numerous efforts in the community toward developing coupled
transport-depletion calculations with OpenMC~\citep{gul2017ane,
lanversin2017icone, lanversin2019phd, liu2019nst, zhuang2020pne, zhao2020ned,
zhao2020cpc, zhang2020ane}. While these efforts have proven the viability of
performing such coupled calculations with OpenMC, none of them has resulted in a
validated, openly-available, and generalized capability that OpenMC users can
take advantage of. The capability that is described herein is part of the OpenMC
project and can be found in version 0.11~\citep{openmc-0110} of the code or
later.

The paper is organized as follows. In \cref{sec:methods}, we describe the
methods and algorithms used to solve the depletion equations, the implementation
within OpenMC, and how nuclear data is used to create depletion chains. Then, in
\cref{sec:results}, we present results on two problems that demonstrate the
accuracy of OpenMC when compared to Serpent~\citep{leppanen2015ane}. In
\cref{sec:conclusions}, we summarize the findings, discuss limitations, and
provide commentary on future avenues of inquiry that may be promising.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec:methods}

The equation that governs the transmutation of nuclides by nuclear reactions and
radioactive decay inside of an irradiated environment can be written as
\begin{equation}
  \label{eq:depletion}
  \begin{split}
  \frac{dN_i(t)}{dt} = &\sum\limits_j \left [ f_{j \rightarrow i}
  \int_0^\infty dE \; \sigma_j (E, t) \phi(E,t) + \lambda_{j\rightarrow i}
  \right ] N_j(t) \\ &- \left [\int_0^\infty dE \; \sigma_i (E,t) \phi(E,t) +
  \sum\limits_j \lambda_{i\rightarrow j} \right ] N_i(t), \quad i=1,\dots,n
  \end{split}
\end{equation}
where
\begin{equation*}
  \begin{split}
  N_i(t) &\equiv \text{density of nuclide $i$ at time $t$} \\
  \sigma_i(E,t) &\equiv \text{transmutation cross section for nuclide $i$ at energy $E$ and time $t$} \\
  \phi(E,t) &\equiv \text{neutron flux at energy $E$ and time $t$} \\
  f_{j \rightarrow i} &\equiv \text{fraction of transmutation reactions in nuclide $j$ that produce nuclide $i$} \\
  \lambda_{j \rightarrow i} &\equiv \text{decay constant for decay modes in nuclide $j$ that produce nuclide $i$} \\
  n &\equiv \text{total number of nuclides.}
  \end{split}
\end{equation*}
Note that we have not included the spatial dependence of the flux or cross
sections. For \cref{eq:depletion} to be accurate, the spatial variation of the
reaction rates over the region which is being depleted should be small. For a
reactor model, this may entail using a unique material in each fuel pin that is
depleted separately. Ultimately, the user is responsible for deciding what level
of spatial discretization is adequate for the problem at hand.

\Cref{eq:depletion} simply states that the rate of change of $N_i$ is equal to
the production rate minus the loss rate. Because the equation for $N_i$ depends
on the density of other nuclides, $\{N_j \,|\, j \ne i\}$, \cref{eq:depletion}
represents a system of first-order ordinary differential equations (ODEs). To
form a proper initial value problem, the nuclide densities at time 0 must be
specified:
\begin{equation}
    N_i(0) = N_{i,0}.
\end{equation}
These equations can be written more compactly in matrix notation as
\begin{equation}
  \label{eq:depletion-matrix-t}
  \frac{d\vect{n}}{dt} = \vect{A}(\vect{n},t)\vect{n}, \quad \vect{n}(0) =
  \vect{n}_0
\end{equation}
where
\begin{equation}
  \vect{n} = \begin{pmatrix} N_1 \\ N_2 \\ \vdots \\ N_n \end{pmatrix}, \quad
  \vect{n}_0 = \begin{pmatrix} N_{1,0} \\ N_{2,0} \\ \vdots \\ N_{n,0} \end{pmatrix}
\end{equation}
and $\vect{A}(\vect{n},t) \in \mathbb{R}^{n\times n}$ is the burnup matrix
containing the decay and transmutation coefficients. Note that the burnup matrix
depends on $\vect{n}$ because the solution to the transport equation depends
on the nuclide densities, thereby making \cref{eq:depletion-matrix-t} nonlinear.

Although the neutron transport equation is time-dependent, the timescale over
which material compositions change is sufficiently long such that the transport
equation can be solved as a time-independent equation. Without loss of
generality, let us write the burnup matrix solely as a function of $\vect{n}$
so that \cref{eq:depletion-matrix-t} becomes
\begin{equation}
  \label{eq:depletion-matrix}
  \frac{d\vect{n}}{dt} = \vect{A}(\vect{n})\vect{n}, \quad \vect{n}(0) =
  \vect{n}_0.
\end{equation}
This better reflects the fact that the time, $t$, is generally not an input to
the transport equation, and hence the burnup matrix is uniquely determined by
the value of $\vect{n}$.

The principle difficulty of solving \cref{eq:depletion-matrix} is the fact that
the elements of $\vect{A}$ are not constant with respect to time. If
$\vect{A}$ is constant, \cref{eq:depletion-matrix} admits a closed-form
solution,
\begin{equation}
  \label{eq:constant-A}
  \vect{n}(t) = \exp \left (\vect{A} t \right ) \vect{n}_0.
\end{equation}
The exponential of a matrix $\vect{X}$ is defined by the power series
expansion
\begin{equation}
  \exp(\vect{X}) = \sum\limits_{k=0}^\infty \frac{\vect{X}^k}{k!}
\end{equation}
where $\vect{X}^0 = \vect{I}$. Solving \cref{eq:depletion-matrix} usually
involves two separate components:
\begin{enumerate}
  \item Use of a numerical method to integrate \cref{eq:depletion-matrix}
  forward in time, usually in a form that involves taking one or more matrix
  exponentials, and
  \item Evaluating the matrix exponential or, equivalently, the action of a
  matrix exponential on a vector.
\end{enumerate}
We will henceforth refer to the first component as \emph{time integration}. Let
us now consider how each of these components are treated in OpenMC.

\subsection{Time Integration}
\label{sec:time_integration}

Because an analytical solution to \cref{eq:depletion-matrix} is not available,
one must resort to using a numerical method that generates approximate solutions
in discrete increments. The problem we need to solve then is: given $\vect{n}_i
\equiv \vect{n}(t_i)$, what is the value of $\vect{n}_{i+1} \equiv
\vect{n}(t_{i+1}) = \vect{n}(t_i + h)$, with $h$ being the step size? Without
loss of generality, we shall choose $t_i=0$ as it makes the presentation of
various methods simpler.

OpenMC doesn't rely on a single time integration method but instead has several
methods available that the user can choose. In the following sections, we
outline the available methods in OpenMC, several of which have not been
previously described in the literature.

\subsubsection{Predictor Method}

The simplest numerical method for solving \cref{eq:depletion-matrix} over a
timestep is known as the \emph{predictor} method. It proceeds by assuming that
$\vect{A}$ is constant over the timestep and assumes the value at the
beginning of the step. The solution then follows directly from
\cref{eq:constant-A}:
\begin{equation}
  \label{eq:predictor}
  \vect{n}_{i+1} = \exp\left(h\vect{A}(\vect{n}_i) \right) \vect{n}_i.
\end{equation}
The end-of-step composition is thus ``predicted'' by the reaction rates at the
beginning of the step. This method has the virtue of requiring a single
transport solution and a single matrix exponential per timestep but is only
first-order accurate.

\subsubsection{predictor--Corrector Methods}

A series of \emph{predictor--corrector} methods that use multiple stages offer
improved accuracy over the predictor method. The simplest of these methods is
the CE/CM algorithm, which is the default algorithm used in
MCNP6~\citep{fensin2006tans}. In this method, $\vect{A}$ is evaluated at the
beginning of the step and used to deplete the composition over half the step.
Then, $\vect{A}$ is reevaluated using the predicted middle-of-step composition
and used to deplete over the full step. The algorithm can be expressed as
\begin{equation}
  \begin{split}
    \vect{n}_{i+1/2} &= \exp \left (\frac{h}{2}\vect{A}(\vect{n}_i) \right) \vect{n}_i \\
    \vect{n}_{i+1} &= \exp \left(h \vect{A}(\vect{n}_{i+1/2}) \right) \vect{n}_i.
  \end{split}
\end{equation}
This method requires two transport solutions and two matrix exponential solves
per timestep (twice as expensive as the predictor method) and is second-order
accurate.

More sophisticated predictor--corrector methods have been introduced by
\citet{isotalo2011ane2,isotalo2011ane3}. In the CE/LI (constant extrapolation,
linear interpolation) method, the end-of-step composition is first predicted
using $\vect{A}(\vect{n}_i)$:
\begin{equation}
  \vect{n}_{i+1}^p = \exp \left( h\vect{A}(\vect{n}_i) \right) \vect{n}_i.
\end{equation}
Then, $\vect{n}_{i+1}^p$ is used to reevaluate $\vect{A}$. An approximation of
$\vect{A}$ that depends on time is formed by linearly interpolating between
$\vect{A}(\vect{n}_i)$ and $\vect{A}\left(\vect{n}_{i+1}^p\right)$.
\Cref{eq:depletion-matrix} thus becomes
\begin{equation}
  \label{eq:linear-ode}
  \frac{d\vect{n}}{dt} = \hat{\vect{A}}(t) \vect{n}
\end{equation}
where
\begin{equation}
  \label{eq:linear-interp}
  \hat{\vect{A}}(t) = \left ( 1 - \frac{t}{h} \right) \vect{A}(\vect{n}_i) +
  \frac{t}{h} \vect{A}\left(\vect{n}_{i+1}^p \right).
\end{equation}
We see that $\hat{\vect{A}}$ no longer depends on $\vect{n}$ thereby making
\cref{eq:linear-ode} a linear nonautonomous ODE. Because $\hat{\vect{A}}$ is a
function of time, an approximation is still needed to obtain $\vect{n}_{i+1}$.
The simplest approach~\citep{isotalo2011ane2} is to average
\cref{eq:linear-interp} over the step,
\begin{equation}
  \bar{\vect{A}} = \int_0^h dt \, \hat{\vect{A}}(t) = \frac{\vect{A}(\vect{n}_i)
  + \vect{A}(\vect{n}_{i+1}^p)}{2},
\end{equation}
and then use $\bar{\vect{A}}$ to deplete over the full step:
\begin{equation}
  \vect{n}_{i+1} = \exp \left ( h \bar{\vect{A}} \right ) \vect{n}_i.
\end{equation}
\citet{isotalo2011ane3} suggested an improvement over this approach wherein
$\hat{\vect{A}}$ is averaged over substeps instead of the entire interval. That
is, if we divide the timestep into $m$ substeps, the solution is given by
\begin{equation}
  \begin{split}
    \label{eq:celi-substeps}
    \vect{A}_s &= \int_{(s-1)h/m}^{sh/m} dt \; \hat{\vect{A}}(t), \quad s=1,\dots,m \\
    \vect{n}_{i+1} &= \exp\left(\vect{A}_m \right) \exp \left(\vect{A}_{m-1}\right)
    \cdots \exp\left(\vect{A}_1\right) \vect{n}_i.
  \end{split}
\end{equation}
While using substeps is a reasonable approach for solving
\cref{eq:linear-ode}, there are other methods as well. For example, some
codes use a backward differentiation formula rather than a method based on an
exponential integrator~\citep{carpenter2009mc,hykes2017mc,sublet2017nds}.

\subsubsection{Magnus expansion integrators}

Any linear approximation of \cref{eq:depletion-matrix} characterized by
$\vect{A}$ that depends only on time, such as \cref{eq:linear-ode}, has an
infinite series solution known as the Magnus expansion~\citep{blanes2009pr}.
This expansion takes the form
\begin{equation}
  \begin{split}
  \vect{n}(t) &= \exp \left( \vect{\Omega}(t) \right ) \vect{n}_0 \\
  \vect{\Omega}(t) &= \sum\limits_{k=1}^\infty \vect{\Omega}_k(t).
  \end{split}
\end{equation}
The first three terms of the expansion are given by
\begin{equation}
  \label{eq:magnus-terms}
  \begin{split}
    \vect{\Omega}_1(t) &= \int_0^t dt_1 \vect{A}(t_1) \\
    \vect{\Omega}_2(t) &= \int_0^t dt_1 \int_0^{t_1} dt_2 [\vect{A}(t_1), \vect{A}(t_2) ] \\
    \vect{\Omega}_3(t) &= \int_0^t dt_1 \int_0^{t_1} dt_2 \int_0^{t_2} dt_3
      \left ( [\vect{A}(t_1), [\vect{A}(t_2), \vect{A}(t_3)]] + [\vect{A}(t_3), [\vect{A}(t_2), \vect{A}(t_1)]] \right )
  \end{split}
\end{equation}
where $[\vect{A},\vect{B}] = \vect{A}\vect{B} - \vect{B}\vect{A}$ is a matrix
commutator. Subsequent terms can be defined through a recurrence relation.

The Magnus expansion can be used as the basis for developing numerical
integration algorithms and has been widely studied. The typical approach is to
truncate the $\vect{\Omega}$ series at an appropriate order and apply an
approximation to the multivariate integrals in \cref{eq:magnus-terms}. In
particular, OpenMC uses a fourth-order commutator-free\footnote{The integrator
is called commutator-free because it does not require computing any matrix
commutators like those that appear in \cref{eq:magnus-terms}. Other methods
based on the Magnus expansion do require computing matrix commutators, some of
which were explored by \citet{josey2017phd}.} exponential
integrator~\citep{blanes2006anm} that is derived from the Magnus expansion in
order to solve \cref{eq:linear-ode}. In this approximation, the Magnus expansion
is truncated at two terms and then expressed as the composition of two
exponentials. For a system of ODEs in the form of \cref{eq:linear-ode}, the
resulting method is
\begin{equation}
  \label{eq:compose4}
  \vect{n}_{i+1} = \exp (a_2 \vect{A}_1 + a_1 \vect{A}_2)
  \exp (a_1 \vect{A}_1 + a_2 \vect{A}_2) \vect{n}_i
\end{equation}
where $\vect{A}_1$ and $\vect{A}_2$ correspond $\hat{\vect{A}}$ evaluated at the
nodes of a two-point Gauss-Legendre quadrature over the interval $[0,h]$:
\begin{equation}
  \begin{split}
    a_1 &= \frac{1}{4} + \frac{\sqrt{3}}{6}, \; a_2 = \frac{1}{4} - \frac{\sqrt{3}}{6} \\
    c_1 &= \frac{1}{2} - \frac{\sqrt{3}}{6}, \; c_2 = \frac{1}{2} + \frac{\sqrt{3}}{6} \\
    \vect{A}_1 &= h\hat{\vect{A}}\left ( c_1 h\right ), \; \vect{A}_2 = h\hat{\vect{A}}\left ( c_2 h\right ).
  \end{split}
\end{equation}
Evaluating $\vect{A}_1$ and $\vect{A}_2$ using \cref{eq:linear-interp} and
substituting in \cref{eq:compose4}, we obtain a CE/LI method based on the
fourth-order commutator-free Magnus integrator:
\begin{equation}
  \label{eq:celi-cfq4}
  \begin{split}
  \vect{n}_{i+1}^p &= \exp \left ( h \vect{A}(\vect{n}_i ) \right ) \\
  \vect{n}_{i+1} &= \exp \left( \frac{h}{12} \vect{A}(\vect{n}_i) +
    \frac{5h}{12} \vect{A}(\vect{n}_{i+1}^p) \right)
  \exp \left( \frac{5h}{12} \vect{A}(\vect{n}_i) +
  \frac{h}{12} \vect{A}(\vect{n}_{i+1}^p) \right) \vect{n}_i.
  \end{split}
\end{equation}
\Cref{eq:celi-cfq4} requires two transport solutions and two matrix
exponentials, whereas the CE/LI method with substeps, \cref{eq:celi-substeps},
requires two transport solutions and $m$ matrix exponentials. A comparison of
these two approaches has shown that \cref{eq:celi-cfq4} achieves better accuracy
than the use of five substeps~\citep{josey2017phd}. OpenMC also includes an
integrator based on the LE/QI (linear extrapolation, quadratic interpolation)
method~\citep{isotalo2011ane2} where the integration of \cref{eq:linear-ode} (in
which case $\hat{\vect{A}}(t)$ is represented by a quadratic polynomial in $t$)
is performed using the fourth-order commutator-free Magnus integrator based on
\cref{eq:compose4}.

\subsubsection{Lie Group Integrator}

OpenMC also includes an integrator based on a commutator-free Lie group
integration method~\citep{celledoni2004fgcs}. In particular, a fourth-order
scheme called CF4 is used that is given as
\begin{equation}
  \begin{split}
    \vect{A}_1 &= h\vect{A}(\vect{n}_0) \\
    \hat{\vect{n}}_1 &= \exp \left ( \frac{\vect{A}_1}{2} \right ) \\
    \vect{A}_2 &= h\vect{A}(\hat{\vect{n}}_1) \\
    \hat{\vect{n}}_2 &= \exp \left ( \frac{\vect{A}_2}{2} \right ) \\
    \vect{A}_3 &= h \vect{A}(\hat{\vect{n}}_2) \\
    \hat{\vect{n}}_3 &= \exp \left ( -\frac{\vect{A}_1}{2}  + \vect{A}_3 \right ) \\
    \vect{A}_4 &= h\vect{A}(\hat{\vect{n}}_3) \\
    \vect{n}_{i+1} &= \exp \left ( \frac{\vect{A}_1}{4} + \frac{\vect{A}_2}{6} + \frac{\vect{A}_3}{6} - \frac{\vect{A}_4}{12} \right )
    \exp \left ( -\frac{\vect{A}_1}{12} + \frac{\vect{A}_2}{6} + \frac{\vect{A}_3}{6} - \frac{\vect{A}_4}{4} \right ) \vect{n}_i.
  \end{split}
\end{equation}
This method requires four transport solutions and five matrix exponentials per
timestep. However, because it is fourth-order, it allows large timesteps to be
used without sacrificing accuracy. For example, on a test problem based on a
3$\times$3 fuel pin array with gadolinum as a burnable absorber, the CF4 method
allowed 20 day timesteps while still achieving an accuracy of of 0.1\% on the
$^{157}$Gd number density~\citep{josey2017phd}.

\subsubsection{Extended Predictor--Corrector Methods}

One option to improve on the traditional predictor--corrector methods is to add
intermediate steps at which $\vect{A}$ is evaluated and combine them in such a
way as to reduce the error of the solution. The general method would take the
following form:
\begin{equation}
  \label{eq:epc}
  \begin{split}
    \hat{\vect{n}}_1 &= \vect{n}_i \\
    \hat{\vect{n}}_k &= \exp \left( h \sum_{j=1}^{k-1} a_{ij} \vect{A}(\hat{\vect{n}}_j) \right) \vect{n}_i \\
    \vect{n}_{i+1} &= \exp \left( h \sum_{j=1}^{s} b_{j} \vect{A}(\hat{\vect{n}}_j) \right) \vect{n}_i
  \end{split}
\end{equation}
Noting that the coefficients in \cref{eq:epc} can be arranged in a Butcher
tableau, \citet{josey2016jcp} proposed using the coefficients for the classic
fourth Runge--Kutta method:
\begin{equation}
  \begin{split}
    \vect{A}_1 &= h\vect{A}(\vect{n}_0) \\
    \hat{\vect{n}}_1 &= \exp \left ( \frac{\vect{A}_1}{2} \right ) \\
    \vect{A}_2 &= h\vect{A}(\hat{\vect{n}}_1) \\
    \hat{\vect{n}}_2 &= \exp \left ( \frac{\vect{A}_2}{2} \right ) \\
    \vect{A}_3 &= h \vect{A}(\hat{\vect{n}}_2) \\
    \hat{\vect{n}}_3 &= \exp \left ( \vect{A}_3 \right ) \\
    \vect{A}_4 &= h\vect{A}(\hat{\vect{n}}_3) \\
    \vect{n}_{i+1} &= \exp \left ( \frac{\vect{A}_1}{6} + \frac{\vect{A}_2}{3}
      + \frac{\vect{A}_3}{3} + \frac{\vect{A}_4}{6} \right ) \vect{n}_i.
  \end{split}
\end{equation}

\subsubsection{Stochastic Implicit Euler Integrators}

All of the integration methods discussed thus far are explicit methods that are
only conditionally stable. There are many well-documented cases where
oscillatory behavior and diverging solutions have been observed. To address
this, \citet{dufek2013ane} proposed an implicit-Euler-like method that is
unconditionally stable. The main idea is to replace $\vect{A}(\vect{n}_i)$ with
$\vect{A}(\vect{n}_{i+1})$ in \cref{eq:predictor}:
\begin{equation}
  \label{eq:implicit}
  \vect{n}_{i=1} = \exp \left( h\vect{A}(\vect{n}_{i+1}) \right) \vect{n}_i.
\end{equation}
This results in a non-linear equation for $\vect{n}_{i+1}$ that can't be solved
directly. However, it is possible to solve for $\vect{n}_{i+1}$ using a
fixed-point iteration. \citet{dufek2013ane} utilized a stochastic approximation
based on the Robbins-Monro algorithm~\citep{robbins1951ams} in order to
guarantee convergence of the fixed-point iteration while accounting for
stochasticity in the MC transport solver.

The major downside of an algorithm based on \cref{eq:implicit} is that, like the
predictor method, it is only first-order accurate. Implicit methods have also
been applied to higher-order
algorithms~\citep{kotlyar2014ane,kotlyar2016ane,cosgrove2020ane}, thereby
obtaining methods that are unconditionally stable and accurate with longer
timesteps. In OpenMC, stochastic implicit versions of the CE/LI and LE/QI
methods have also been implemented. The basis for the algorithm is to iterate on
$\vect{n}_{i+1}^p$ used in \cref{eq:linear-interp}, obtaining progressively
better estimates that are then used to solve \cref{eq:linear-ode}. A full
discussion of the algorithm and its performance and tradeoffs can be found in
\citep{josey2017phd}.

\subsubsection{Summary}

OpenMC does not rely on a single time integration method but rather has several
classes that implement different algorithms. Generally, there is a tradeoff
between the accuracy of the method, the computational expense, and the memory
requirements. When a MC code is used, the expense is dominated by the time
to compute the transport solution, i.e., to evaluate $\vect{A}$ for a given
$\vect{n}$. Thus, the cost of a method scales with the number of $\vect{A}$
evaluations that are performed per timestep. On the other hand, methods that
require more evaluations generally achieve higher accuracy. The predictor method
only requires one evaluation and its error converges as $\mathcal{O}(h)$. The
CE/CM method requires two evaluations and is thus twice as expensive as the
predictor method, but achieves an error of $\mathcal{O}(h^2)$. An exhaustive
description of time integration methods and their merits can be found in the
thesis of \citet{josey2017phd}.

\subsection{Matrix Exponential}

As we saw in \cref{sec:time_integration}, numerically integrating
\cref{eq:depletion-matrix} requires evaluating one or more matrix exponentials.
OpenMC uses the Chebyshev rational approximation method (CRAM), which was
introduced in a series of papers by Pusa~\citep{pusa2010nse,pusa2011nse}, to
evaluate matrix exponentials. In particular, OpenMC utilizes an incomplete
partial fraction (IPF) form~\citep{pusa2016nse} of CRAM that provides a good
balance of numerical stability and efficiency. In this representation the matrix
exponential is approximated as
\begin{equation}
    \exp(\vect{A}t) \approx \alpha_0 \prod\limits_{\ell=1}^{k/2} \left (
    \vect{I} + 2 \text{Re} \left ( \widetilde{\alpha}_\ell \left (\vect{A}t
    - \theta_\ell \vect{I} \right )^{-1} \right ) \right )
\end{equation}
where $k$ is the order of the approximation and $\alpha_0$,
$\widetilde{\alpha}_\ell$, and $\theta_\ell$ are coefficients that have been
tabulated for orders up to $k=48$. Rather than computing the full approximation
and then multiplying it by a vector, the following algorithm\footnote{The
original description of the algorithm presented by \citet{pusa2016nse} contains
a typo.} is used to incrementally apply the terms within the product:
\begin{algorithm}[H]
  \caption{Incomplete partial fraction form of CRAM.}
  \label{alg:cram}
  \begin{algorithmic}[1]
    \State $\vect{n} \gets \vect{n_0}$
    \For{$\ell \gets 1$ \textbf{to} $k/2$}
      \State $\vect{n} \gets \vect{n} + 2\text{Re}(\widetilde{\alpha}_\ell
        (\vect{A}t - \theta_\ell)^{-1})\vect{n}$
    \EndFor
    \State $\vect{n} \gets \alpha_0 \vect{n}$
  \end{algorithmic}
\end{algorithm}
The $k$th order approximation for CRAM requires solving $k/2$ sparse linear
systems. OpenMC relies on functionality from scipy.sparse.linalg for solving the
linear systems.

\subsection{Implementation}

In-memory handling through Python API. This allows time loading nuclear data to
be amortized since it only needs to be done once.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

\subsection{Data Preparation}

Using ENDF/B-VII.1 files from MCNP distribution. However, for light water
$S(\alpha,\beta)$, Serpent is not capable of handling continuous format, so the
file from ENDF/B-VII.0 was used. The evaluation didn't change between VII.0 and
VII.1, but the processing was different. For Serpent, decay and neutron fission
yield sublibraries are concatenated.

\subsection{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}

This research was supported by the Exascale Computing Project (17-SC-20-SC), a
collaborative effort of the U.S. Department of Energy Office of Science and the
National Nuclear Security Administration. The submitted manuscript has been
created by UChicago Argonne, LLC, operator of Argonne National Laboratory under
contract DE-AC02-06CH11357.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section*{References}

\bibliographystyle{elsarticle-harv}
\bibliography{references}

\clearpage
\vspace*{\fill}
\noindent\fbox{%
  \parbox{\textwidth}{%
    The submitted manuscript has been created by UChicago Argonne, LLC, Operator
    of Argonne \mbox{National} Laboratory (``Argonne'').  Argonne, a
    U.S. Department of Energy Office of Science laboratory, is operated
    under Contract No. \mbox{DE-AC02-06CH11357}.  The U.S. Government retains for
    itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable
    worldwide license in said article to reproduce, prepare derivative works,
    distribute copies to the public, and perform publicly and display publicly,
    by or on behalf of the Government. The Department of Energy will provide
    public access to these results of federally sponsored research in accordance
    with the DOE Public Access
    Plan. \url{http://energy.gov/downloads/doe-public-access-plan.}
  }%
}
\vspace*{\fill}

\end{document}